\documentclass{article}
\usepackage[utf8]{inputenc}


\usepackage{amssymb,amsmath,amsthm}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{breqn}


\title{RIWA changes}
\author{Chris Cremer}
\date{June 2017}

\begin{document}

\maketitle

\section{Changes}




\subsection{Defining the implicit distribution \texorpdfstring{$q_{IW}$}{}}


*Add tilde because its unormalized\\
*IWAE uses eqn 3. Eqn 4 is more\\ accurate version. so its an upper bound
* Add tilde to marginal as well\\
* Remove the comment about batches\\
*Equation (4)  What are we\\ marginalizing out? Remove it? \\




In this section, we derive the implicit distribution that arises from importance sampling from a distribution $p$ using $q$ as a proposal distribution.

Given a batch of samples $z_{1}...z_{k}$ from $q(z|x)$, the following is the unnormalized importance weighted $q_{IW}$ distribution as a function of one of the samples, $z_i$:
\begin{align} 
    \tilde{q}_{IW}(z_i|x,z_{\setminus i}) = k \tilde{w}_i q(z_i|x)
    &= \left( \frac{ \frac{p(x,z_i)}{q(z_i|x)}}{  \frac{1}{k}   \sum_{j=1}^k \frac{p(x,z_j)}{q(z_j|x)}}  \right) q(z_i|x) \\
    &= \frac{p(x,z_i)}{\frac{1}{k} \left(  \frac{p(x,z_i)}{q(z_i|x)}+ \sum_{j=2}^k \frac{p(x,z_j)}{q(z_j|x)} \right)} 
\end{align}
For the equation above, we are arbitrarily setting $i$ to be the first index and $j$ is used to index all the other $k-1$ samples. Here are some properties to note of $\tilde{q}_{IW}$:  when $k=1$, $q_{IW}(z|x)$ will be equal to $q(z|x)$, when $k > 1$, we see that the form of $q_{IW}$ depends on the true posterior $p$, and when $k=\infty$, $q_{IW}(z|x)$ becomes the true posterior $p(z|x)$. See the appendix for details.

The marginal distribution $q_{IW}(z|x)$ is given by:
\begin{align} 
    q_{IW}(z|x) &= E_{z_{2}...z_{k} \sim q(\cdot |x)} \left[ \frac{p(x,z)}{  \frac{1}{k} \left( \frac{p(x,z)}{q(z|x)}+ \sum_{j=2}^k \frac{p(x,z_j)}{q(z_j|x)} \right) } \right] \label{marg} 
\end{align}

Figure \ref{viz1} visualizes $q_{IW}$ on a 2D distribution approximation problem. The base distribution $q$ is a Gaussian.  As we increase the number of samples $k$ used for the sampling-resampling, the approximation approaches the true distribution. This distribution is nonparametric in the sense that, as the true posterior grows more complex, so does the shape of $q_{IW}$.






\subsection{Recovering the IWAE bound from the VAE bound}

If we now set $q(z|x) = q_{IW}(z|x)$, then we recover the IWAE ELBO:
\begin{align}
        \mathcal{L}_{VAE}[q_{IW}]
        &= E_{z_{1}...z_{k} \sim q_{IW}(z|x)} \left[  \frac{1}{k}\sum_{i=1}^k log\left(\frac{p(x,z_i)}{q_{IW}(z_i|x,z_{\setminus i})}  \right)  \right] \\
    &= E_{z_{1}...z_{k} \sim q(z|x)} \left[ \sum_{l=1}^k \tilde w_l \frac{1}{k}\sum_{i=1}^k log\left(\frac{p(x,z_i)}{\frac{p(x,z_i)}{\frac{1}{k}   \sum_{j=1}^k \frac{p(x,z_j)}{q(z_j|x)}}}  \right)  \right] \\
    %&= E_{z_{1}...z_{k} \sim (z|x)} \left[ \frac{1}{k}\sum_{i=1}^k log\left(\frac{1}{k} \sum_{j=1}^k \frac{p(x,z_j)}{q(z_j|x)}\right)  \right] \label{with_avg} \\   
    &= E_{z_{1}...z_{k} \sim q(z|x)} \left[  log\left(\frac{1}{k}\sum_{j=1}^k \frac{p(x,z_j)}{q(z_j|x)}  \right)  \right] = \mathcal{L}_{IWAE}[q]
\end{align}









\subsection{Discussion}
Add this to discussion:\\

\begin{align} 
    log(p(x)) \geq L_{IWAE}[q_{IW}] = L_{VAE}[q_{IW}] \geq L_{IWAE}[q] = L_{VAE}[q_{IW|z\setminus i}] \geq L_{VAE}[q] \nonumber
\end{align}



\subsection{Acknowledgements}
We'd also like to thank Yuri Burda, Christian Naesseth, and Scott Linderman for bringing our attention to oversights in the paper.


\end{document}
